---
title: "Getting Started with gstIndia"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with gstIndia}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment="#>", fig.width=7, fig.height=4.5)
```

**gstIndia** v2.0.0 provides nine tidy `data.table` datasets covering India's
GST ecosystem from July 2017 through early 2026, along with a complete suite of
analysis functions. All monetary values are in **Rs. Crore**.

```{r load}
library(gstIndia)
library(data.table)
```

---

## What's in the package?

```{r catalogue}
gst_catalogue()
```

---

## Quick reference

```{r ref}
gst_years()      # available financial years
gst_states()     # 36+ states with codes and regions
```

---

## 1. State-wise Collection

### Filter and explore

```{r filter}
# South India FY 2023-24, total column only
gst_filter(region = "South", financial_year = "2023-24", component = "total")
```

### Top states

```{r top}
gst_top_states(n = 10, financial_year = "2023-24")
```

### Annual summary

```{r annual}
# National totals per FY
gst_annual_summary(by = "national")[, .(financial_year, total)]
```

### State share

```{r share}
gst_state_share(financial_year = "2023-24")[1:8]
```

### Year-on-year growth

```{r yoy}
gst_yoy(state = "Maharashtra")[order(-month_date)][1:6,
  .(month, current, prior, pct_growth)]
```

### Component mix

```{r mix}
gst_component_mix(state = "Gujarat", financial_year = "2023-24")
```

---

## 2. Refunds & Net Collection

```{r net}
# Net after refunds — top 10 states FY 2023-24
gst_net_collection(financial_year = "2023-24", by = "state")[order(-net)][1:10]
```

---

## 3. Return Filing Compliance

```{r compliance}
# GSTR-3B average compliance by region FY 2023-24
gst_compliance_trend("GSTR-3B", by = "region", financial_year = "2023-24")[
  order(-avg_filing_pct)]
```

```{r low-compliance}
# States below 90% GSTR-1 compliance in FY 2022-23
gst_low_compliance_states("GSTR-1", threshold = 0.90, financial_year = "2022-23")
```

---

## 4. E-Way Bills

```{r ewb}
# Top 5 states by intrastate EWB count FY 2023-24
ewb_top_states(n = 5, financial_year = "2023-24", direction = "intrastate")
```

```{r ewb-value}
# Top 5 by interstate outgoing assessable value
ewb_top_states(n = 5, metric = "assessable_value",
               direction = "interstate_out", financial_year = "2023-24")
```

---

## 5. IGST Settlement to States

```{r igst}
igst_settlement_summary(financial_year = "2023-24", by = "state")[order(-total)][1:8]
```

---

## 6. Taxpayer Registration

```{r reg}
# Top 10 states by normal taxpayer registrations
gst_registration_summary()[order(-normal)][1:10]
```

```{r profile}
# Taxpayer constitution profile
gst_taxpayer_profile[order(-total_taxpayers)]
```

---

## 7. Gross vs Net National Collection (FY 2024-26)

```{r gross-net}
gross_net_collection[, .(month, gross_domestic_cur, refund_cur)]
```

---

## Example: ggplot2 Trend Chart

```{r plot, eval=FALSE}
library(ggplot2)

top5 <- gst_top_states(n = 5, financial_year = "2023-24")$state
dt   <- gst_filter(state = top5,
                   financial_year = c("2022-23","2023-24","2024-25"),
                   component = "total")

ggplot(dt, aes(month_date, total / 1e3, colour = state)) +
  geom_line(linewidth = 0.8) +
  scale_y_continuous(labels = scales::comma_format(suffix = "K Cr")) +
  labs(title    = "Monthly GST Collection — Top 5 States",
       subtitle = "FY 2022-23 to 2024-25  |  Rs. '000 Crore",
       x = NULL, y = NULL, colour = NULL,
       caption  = "Source: Ministry of Finance, GoI") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```

---

## Data Source

All data sourced from the **Ministry of Finance, Government of India** and the
**GST Council**. Data covers July 2017 (GST rollout) through February 2026.
FY 2025-26 figures are provisional.
