---
title: "Getting Started with gstIndia"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with gstIndia}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.width = 7,
  fig.height = 4
)
```

## Overview

**gstIndia** ships two ready-to-use `data.table` datasets covering India's GST
collections from July 2017 through early 2026, along with a suite of helper
functions for common analytical tasks.

```{r load}
library(gstIndia)
library(data.table)
```

---

## Datasets at a Glance

### `gst_statewise` — Monthly state/UT breakdown

```{r statewise-preview}
data(gst_statewise)
gst_statewise
```

Key columns: `financial_year`, `month`, `month_date`, `state_code`, `state`,
`region`, `cgst`, `sgst`, `igst`, `cess`, `total` (all monetary values in
**Rs. Crore**).

### `gst_allindia` — All-India monthly summary

```{r allindia-preview}
data(gst_allindia)
gst_allindia
```

Available categories: `domestic`, `igst_import`, `gross_revenue`, `refund`,
`net_revenue`.

---

## Helper Functions

### `gst_years()` and `gst_states()`

```{r reference}
gst_years()
gst_states()
```

### `gst_filter()` — Subset the data

```{r filter}
# FY 2023-24, Southern states, total only
gst_filter(
  financial_year = "2023-24",
  region         = "South",
  component      = "total"
)
```

### `gst_annual_summary()` — Annual aggregates

```{r annual}
# Annual state-level totals for FY 2022-23 and 2023-24
gst_annual_summary(
  by             = "state",
  financial_year = c("2022-23", "2023-24"),
  component      = c("cgst", "sgst", "igst", "total")
)[order(-total)][1:10]
```

### `gst_top_states()` — Ranking

```{r top-states}
gst_top_states(n = 10, financial_year = "2023-24")
```

### `gst_state_share()` — Share of national total

```{r state-share}
share <- gst_state_share(financial_year = "2023-24")
share[order(-share_pct)][1:10]
```

### `gst_yoy()` — Year-on-year growth

```{r yoy}
yoy <- gst_yoy(state = "Maharashtra")
yoy[order(-month_date)][1:12, .(month, current, prior, pct_growth)]
```

### `gst_component_mix()` — Component share over time

```{r comp-mix}
gst_component_mix(state = "Gujarat", financial_year = "2023-24")
```

---

## Example Analysis with ggplot2

```{r plot-top5, eval=FALSE}
library(ggplot2)

# Monthly trend — top 5 states FY 2023-24
top5 <- gst_top_states(n = 5, financial_year = "2023-24")$state
dt   <- gst_filter(state = top5, financial_year = c("2022-23","2023-24"),
                   component = "total")

ggplot(dt, aes(x = month_date, y = total / 1e3, colour = state)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 1.5) +
  scale_y_continuous(labels = scales::comma_format(suffix = " K Cr")) +
  labs(
    title    = "Monthly GST Collection — Top 5 States",
    subtitle = "FY 2022-23 to 2023-24  |  Rs. '000 Crore",
    x = NULL, y = NULL, colour = "State"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```

---

## Data Source

All data is sourced from the **Ministry of Finance, Government of India** — GST
Revenue Collection releases published on <https://www.gst.gov.in>.

Values are in **Rs. Crore** (1 Crore = 10 million Indian Rupees).
FY 2017-18 covers only July 2017 – March 2018 (GST was rolled out on
1 July 2017). FY 2025-26 data is provisional (up to February 2026).
